---
title: "Development Vignette"
autor: "Mathis Erler"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Development Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
This markdown shows the current state of the MCMC extension of the USE package.

In a first step we load the needed packages.
```{r setup, message=FALSE}
devtools::install_github("mdmdma/USE.MCMC")
library(USE.MCMC)
library(terra)
library(virtualspecies)
library(sf)
library(ggplot2)
library(parallel)
library(FNN)
```
As a first step we fix a seed to get reproducible results.
```{r fix.seed}
seed.number = 43
set.seed(seed.number)
```

In a next step we can load data. For conveniance the worldclim dataset of the USE package can be used. It contains 5 variables at a resolution of 10 arcminutes.
```{r data}
env.data.raster <- USE.MCMC::Worldclim_tmp %>%
  terra::rast( type="xyz")
```
If available other compibations of variables can be used
```{r data.if.available, eval=FALSE}
data.dir <- "/home/mathis/Desktop/semesterarbeit10/data"

number.of.dimensions <- 5
env.data.raster.all.dim <- geodata::worldclim_global(var='bio', res=10, path=data.dir)  %>%
  terra::crop(terra::ext(-12, 25, 36, 60))

selected.layers <- sample(terra::nlyr(env.data.raster.all.dim), number.of.dimensions)
env.data.raster <- env.data.raster.all.dim[[selected.layers]]
```
The data is further prepared to be later used to plot the distribution of the environment.
```{r prepare.data}
env.data.sf <- env.data.raster %>%
  as.data.frame(xy = TRUE) %>%
  sf::st_as_sf(coords = c("x", "y"))

# Generate the environmental space using PCA
rpc <- rastPCA(env.data.raster,  stand = TRUE)
raster.with.pc <- c(env.data.raster, rpc$PCs)

# Attaching the data in the PCA coordinates
env.with.pc.fs <- rpc$PCs %>%
  as.data.frame(xy = TRUE) %>%
  na.omit() %>%
  sf::st_as_sf(coords = c("x", "y")) %>%
  st_join(env.data.sf)
```
We sample a virtual species analog to the method from the USE vignette.
```{r create.vs, message=FALSE}
virtual.presence.data <- getVirtualSpeciesPresencePoints(
  env.data = env.data.raster, n.samples = 300)
virtual.presence.points <- virtual.presence.data$sample.points
virtual.presence.points.pc <- terra::extract(raster.with.pc,
                                             virtual.presence.points, bind = TRUE) %>%
  sf::st_as_sf()
```
Now that the data and species are prepared, we can sample pseudo absences. Here we will compare multiple methods.

First we sample using the method presented in the paper. 
```{r uniform.paper}
sampled.points.uniform.p <- paSampling(env.rast = env.data.raster, pres = virtual.presence.points, grid.res = 10)
sampled.points.uniform.p.location <- sf::st_drop_geometry(sampled.points.uniform.p) %>%
  dplyr::select(c("x", "y")) %>%
  sf::st_as_sf(coords = c("x", "y"))
sampled.points.uniform.paper <- terra::extract(env.data.raster.with.pc, sampled.points.uniform.p.location)
```
The sampled points were rearanged so that the data can be easier compared. 

As a second comparison I implemented an other grid based uniform sampling approach, see \code{\link{paSamplingNn}} for details. It is similar to the MCMC approach that locations in the PCA space are sampled and then remapped to points in the environmental space using nearest neighbor search. It conceptually forms a bridge to the MCMC approach.
Using the parameter \code{n.sample.points} we can select the number of sampled points.  
```{r uniform.nn}
n.sample.points = 300
sampled.points.uniform.nn <- paSamplingNn(env.rast = env.data.raster,
                                          pres = virtual.presence.points,
                                          grid.res = 15, n.tr = 2,
                                          n.samples = n.sample.points)

```
Analog to the methods shown before we can use the MCMC approach to sample pseudo absences. It works with the same minimal parameters \code{env.rast} and \code{pres}, but can be supplied with additional parameters to controls its behavior.
```{r uniform.mcmc}
sampled.points.mcmc <- paSamplingMcmc(env.data.raster = env.data.raster,
                                      pres = virtual.presence.points, precomputed.pca = rpc, environmental.cutof.percentile = 0.001,
                                      num.chains = 2,
                                      num.cores = 2,
                                      chain.length = 10000,
                                      n.samples = n.sample.points,
                                      verbose = TRUE)
```
We can now analyse the distribution of PC scores of the points sampled by different methods. 
```{r visualize, fig.width=7, fig.height=8, echo=FALSE}
# Create a layout with 3 rows: 2 for plots, 1 for legend
layout(matrix(c(1, 2, 3), nrow = 3, byrow = TRUE), 
       heights = c(2,2,1))  # Make legend row shorter

par(mar = c(4, 4, 5, 2))  # bottom, left, top, right

plots <- invisible(lapply(c("PC1", "PC2", "PC3"), function(col) {
  # Your existing plot code without the legend
  x_range <- range(
    c(density(env.with.pc.fs[[col]])$x,
      density(virtual.presence.points.pc[[col]])$x,
      density(sampled.points.uniform.nn[[col]])$x,
      density(sampled.points.mcmc[[col]])$x)
  )
  y_range <- range(
    c(density(env.with.pc.fs[[col]])$y,
      density(virtual.presence.points.pc[[col]])$y,
      density(sampled.points.uniform.nn[[col]])$y,
      density(sampled.points.mcmc[[col]])$y)
  )
  
  plot(density(env.with.pc.fs[[col]]),
       col = "green",
       main = paste("Density Comparison for", col),
       xlim = x_range,
       ylim = y_range,
       lwd = 2)
  
  lines(density(virtual.presence.points.pc[[col]]), col = "black", lwd = 2)
  lines(density(sampled.points.uniform.paper[[col]]), col = "orange", lwd = 2)
  lines(density(sampled.points.uniform.nn[[col]]), col = "blue", lwd = 2)
  lines(density(sampled.points.mcmc[[col]]), col = "red", lwd = 2)
}))

# Create an empty plot for the legend
par(mar = c(0, 0, 0, 0))  # Remove margins
plot.new()
legend("center",
       legend = c("Environment", "Virtual Presence", "Sampled Points uniform", 
                  "Sampled Points uniform nn", "Sampled Points mcmc"),
       col = c("green", "black", "orange", "blue", "red"),
       lwd = 2,
       horiz = FALSE,  # Vertical legend (rows)
       bty = "n",      # No box around legend
       cex = 1.5)      # incease text size
```







