---
title: "Development Vignette"
autor: "Mathis Erler"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Development Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
This markdown shows the current state of the MCMC extension of the USE package.

In a first step we load the needed packages.
```{r setup, message=FALSE}
#devtools::install_github("mdmdma/USE.MCMC")
library(USE.MCMC)
library(terra)
library(virtualspecies)
library(sf)
library(ggplot2)
library(parallel)
library(FNN)
```
As a first step we fix a seed to get reproducible results.
```{r fix.seed}
seed.number = 42
set.seed(seed.number)
```

In a next step we can load data. For conveniance the worldclim dataset of the USE package can be used. It contains 5 variables at a resolution of 10 arcminutes.
```{r data}
env.data.raster <- USE.MCMC::Worldclim_tmp %>%
  terra::rast( type="xyz")
```
If available other compibations of variables can be used
```{r data.if.available, eval=FALSE}
data.dir <- "/home/mathis/Desktop/semesterarbeit10/data"

number.of.dimensions <- 5
env.data.raster.all.dim <- geodata::worldclim_global(var='bio', res=10, path=data.dir)  %>%
  terra::crop(terra::ext(-12, 25, 36, 60))

selected.layers <- sample(terra::nlyr(env.data.raster.all.dim), number.of.dimensions)
env.data.raster <- env.data.raster.all.dim[[selected.layers]]
```
The data is further prepared to be later used to plot the distribution of the environment.
```{r prepare.data}
env.data.sf <- env.data.raster %>%
  as.data.frame(xy = TRUE) %>%
  sf::st_as_sf(coords = c("x", "y"))

# Generate the environmental space using PCA
rpc <- rastPCA(env.data.raster,  stand = TRUE)
# Create a raster with the environmental variables as well as pcs
env.data.raster.with.pc <- c(env.data.raster, rpc$PCs)

# Attaching the data in the PCA coordinates
env.with.pc.fs <- rpc$PCs %>%
  as.data.frame(xy = TRUE) %>%
  na.omit() %>%
  sf::st_as_sf(coords = c("x", "y")) %>%
  st_join(env.data.sf)
```
We sample a virtual species analog to the method from the USE vignette.
```{r create.vs, message=FALSE}
virtual.presence.data <- getVirtualSpeciesPresencePoints(
  env.data = env.data.raster.with.pc, n.samples = 300)
virtual.presence.points <- virtual.presence.data$sample.points
virtual.presence.points.pc <- terra::extract(env.data.raster.with.pc,
                                             virtual.presence.points, bind = TRUE) %>%
  sf::st_as_sf()
```
Now that the data and species are prepared, we can sample pseudo absences. Here we will compare multiple methods.

First we sample using the method presented in the paper. 
```{r uniform.paper}
sampled.points.uniform.p <- paSampling(env.rast = env.data.raster, pres = virtual.presence.points, grid.res = 10)
sampled.points.uniform.p.location <- sf::st_drop_geometry(sampled.points.uniform.p) %>%
  dplyr::select(c("x", "y")) %>%
  sf::st_as_sf(coords = c("x", "y"))
sampled.points.uniform.paper <- terra::extract(env.data.raster.with.pc, sampled.points.uniform.p.location)
```
The sampled points were rearanged so that the data can be easier compared. 

As a second comparison I implemented an other grid based uniform sampling approach, see \code{\link{paSamplingNn}} for details. It is similar to the MCMC approach that locations in the PCA space are sampled and then remapped to points in the environmental space using nearest neighbor search. It conceptually forms a bridge to the MCMC approach.
Using the parameter \code{n.sample.points} we can select the number of sampled points.  
```{r uniform.nn}
n.sample.points = 300
sampled.points.uniform.nn <- paSamplingNn(env.rast = env.data.raster,
                                          pres = virtual.presence.points,
                                          grid.res = 15, n.tr = 2,
                                          n.samples = n.sample.points)

```
Analog to the methods shown before we can use the MCMC approach to sample pseudo absences. It works with the same minimal parameters \code{env.rast} and \code{pres}, but can be supplied with additional parameters to controls its behavior. If verbose is set to true and multiple chains are selected, the output is not usefull.
```{r uniform.mcmc, message=FALSE}
sampled.points.mcmc <- paSamplingMcmc(env.data.raster = env.data.raster,
                                      pres = virtual.presence.points, precomputed.pca = rpc, environmental.cutof.percentile = 0.001,
                                      num.chains = 4,
                                      num.cores = 4,
                                      chain.length = 10000,
                                      n.samples = n.sample.points,
                                      verbose = TRUE)
```
We can now analyse the distribution of PC scores of the points sampled by different methods.

Non of the algorithms try to sample the third dimension uniformly. We can see that the MCMC approach evades the presence stronger than the other approaches, especcialy visible in the second dimension. Importantly we see that on the high end it does not manage to sample points, even tough the species is not present any more. 
```{r visualize, fig.width=7, fig.height=11, echo=FALSE}
# Create a layout with 3 rows: 2 for plots, 1 for legend
if(TRUE) {
  # Create a layout with 3 rows: 2 for plots, 1 for legend
  plotting.dimensions <- c("PC1", "PC2", "PC3")
  layout(matrix(c(1:(length(plotting.dimensions) + 1)), nrow = length(plotting.dimensions) + 1, byrow = TRUE),
         heights = c(rep(3, length(plotting.dimensions)) ,1))  # Make legend row shorter

  par(mar = c(4, 4, 5, 2))  # bottom, left, top, right

  plots <- invisible(lapply(plotting.dimensions, function(col) {
    x_range <- range(
      c(density(env.with.pc.fs[[col]])$x,
        density(sampled.points.uniform.nn[[col]])$x,
        density(sampled.points.mcmc[[col]])$x)
    )
    y_range <- range(
      c(density(env.with.pc.fs[[col]])$y,
        density(sampled.points.uniform.nn[[col]])$y,
        density(sampled.points.mcmc[[col]])$y)
    )

    plot(density(env.with.pc.fs[[col]]),
         col = "green",
         main = paste("Density Comparison for", col),
         xlim = x_range,
         ylim = y_range,
         lwd = 2)

    lines(density(virtual.presence.points.pc[[col]]), col = "black", lwd = 2)
    lines(density(sampled.points.uniform.paper[[col]]), col = "orange", lwd = 2)
    lines(density(sampled.points.uniform.nn[[col]]), col = "blue", lwd = 2)
    lines(density(sampled.points.mcmc[[col]]), col = "red", lwd = 2)
  }))

  # Create an empty plot for the legend
  par(mar = c(0, 0, 0, 0))  # Remove margins
  plot.new()
  legend("center",
         legend = c("Environment", "Virtual Presence", "Sampled Points uniform",
                    "Sampled Points uniform nn", "Sampled Points mcmc"),
         col = c("green", "black", "orange", "blue", "red"),
         lwd = 2,
         horiz = FALSE,  # Vertical legend (rows)
         bty = "n",      # No box around legend
         cex = 1.5)      # incease text size
}
```
In difference to the approaches on which the MCMC approach is based it works in higher dimensions.

Lets explore the changes we have to to to sample in more dimensions.
```{r, uniform.mcmc.higher.dim, message=FALSE}
sampled.points.mcmc.higher.dim <- paSamplingMcmc(env.data.raster = env.data.raster,
                                      pres = virtual.presence.points, precomputed.pca = rpc, environmental.cutof.percentile = 0.001,
                                      num.chains = 4,
                                      num.cores = 4,
                                      chain.length = 10000,
                                      n.samples = n.sample.points,
                                      verbose = TRUE,
                                      dimensions = c("PC1", "PC2", "PC3"))
```
The interpret ability of the results now gets more complicated. We see that the uniform approach now samples the third dimension notably wider, but shows less uniform behavior in the first two dimensions. If this represents the actual environmental space or is an issue with the code itself is unclear, as the density in one dimension depends on the range of possible values in the other dimensions at a given value of the target dimension. It would be a good idea to get insight into this in 3 dimensions to get a idea. This should be 
```{r visualize.higher.dim, fig.width=7, fig.height=11, echo=FALSE}
# Create a layout with 3 rows: 2 for plots, 1 for legend
if(TRUE) {
  # Create a layout with 3 rows: 2 for plots, 1 for legend
  plotting.dimensions <- c("PC1", "PC2", "PC3")
  layout(matrix(c(1:(length(plotting.dimensions) + 1)), nrow = length(plotting.dimensions) + 1, byrow = TRUE),
         heights = c(rep(2, length(plotting.dimensions)) ,1))  # Make legend row shorter

  par(mar = c(4, 4, 5, 2))  # bottom, left, top, right

  plots <- invisible(lapply(plotting.dimensions, function(col) {
    x_range <- range(
      c(density(env.with.pc.fs[[col]])$x,
        density(sampled.points.uniform.nn[[col]])$x,
        density(sampled.points.mcmc.higher.dim[[col]])$x)
    )
    y_range <- range(
      c(density(env.with.pc.fs[[col]])$y,
        density(sampled.points.uniform.nn[[col]])$y,
        density(sampled.points.mcmc.higher.dim[[col]])$y)
    )

    plot(density(env.with.pc.fs[[col]]),
         col = "green",
         main = paste("Density Comparison for", col),
         xlim = x_range,
         ylim = y_range,
         lwd = 2)

    lines(density(virtual.presence.points.pc[[col]]), col = "black", lwd = 2)
    lines(density(sampled.points.uniform.paper[[col]]), col = "orange", lwd = 2)
    lines(density(sampled.points.uniform.nn[[col]]), col = "blue", lwd = 2)
    lines(density(sampled.points.mcmc.higher.dim[[col]]), col = "red", lwd = 2)
  }))

  # Create an empty plot for the legend
  par(mar = c(0, 0, 0, 0))  # Remove margins
  plot.new()
  legend("center",
         legend = c("Environment", "Virtual Presence", "Sampled Points uniform",
                    "Sampled Points uniform nn", "Sampled Points mcmc"),
         col = c("green", "black", "orange", "blue", "red"),
         lwd = 2,
         horiz = FALSE,  # Vertical legend (rows)
         bty = "n",      # No box around legend
         cex = 1.5)      # incease text size
}
```






